---
title: "IDM_project"
author: "Kwaku"
date: "11/7/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nimble)
library(coda)
library(mcmcplots)
library(MCMCglmm)
```

The following are used for the simulation:
We assume n.genus  $g = 1,2,...,n.genus$.
* For each g genera, we assume n.species in that geunus
* We assume the data was collected at n.sites
* We also assume the total number of visits to each site is n.visit.
* We also assume that some of the data are collected with species ID
* Most of them are collected with genus ID
* shan.index is the shannon's index
* p.tag is the detectiion probability for the occurence model (species data)
* includecov (T/F) indicates whether it is a model with a covariate (TRUE) or no covariate (FALSE)


```{r include=FALSE}
source("sim_rep.R") #simulation of data
source('nimble.R') #code for the nimble
source('estimation.R') #parametrisation for the NIMBLE
source('shanestimate.R') #Function that estimates the shannon index from the MCMC
```

# Simulating the data
```{r}
set.seed(2020)
includecov= FALSE 
#includecov=TRUE means that there is covariate effect and false otherwise
n.sites=20
n.species = 5
n.visit=3
covte = covariates(n.sites)
data = sim(n.sites, n.species,n.visit,n.sites ,n.sites, cov=covte, includecov) #covte is the covariates
summary(data)
```

# Original Parameters
$$
\begin{split}
\sigma^2_{\alpha} &= 0.4^2\\
\sigma^2_{species} &=\sigma^2_{\gamma} =2\\
\sigma^2_{site} &= \sigma^2_{\beta}=1\\
p.tag &= 0.7
\end{split}
$$


# Running the model
```{r}
mcmc.out <- est(data, code)
```

#Diagnostics of the model
```{r}
 coda_samples <- mcmc(mcmc.out$mcmc.out$samples)
#mcmcplot(coda_samples)
```

# Parameter Estimation

* Mean bias of shannon index

* MSE of shannon index

* Estimate of $\sigma_{\alpha}$, $\sigma_{\beta}$, $\sigma_{\gamma}$ and $p.tag$

```{r}
true_shan <- data$shan.index #true estimates of shannon index from the simulation
if(includecov==TRUE){
  est_beta <- matrix(mcmc.out$mode.out[1:(n.species*3)],
                     nrow=n.species, ncol=3) #estimates of beta from mcmc
  est_shan = shanest(est_beta,data$covariates, TRUE) #estimation of shannon index
  mse_shan = sum(((true_shan-est_shan)^2)/n.sites) #MSE of shannon index
  mbias_shan = mean(true_shan-est_shan) #mean bias of shannon index
   p_est <- mcmc.out$mode.out[(n.species*3)+1] #detection probability
   sig_alpha <- NA
   sig_beta <- NA
   sig_gamma <- NA
  }else{
    est_shan <- mcmc.out$mode.out[2:(n.sites+1)] #estimates of shannon index
    mse_shan = sum(((true_shan-est_shan)^2)/n.sites) #MSE of shannon index
    mbias_shan = mean(true_shan-est_shan)#mean bias of shannon index
  p_est <- mcmc.out$mode.out[1] #detection probability
  sig_alpha <- tail(mcmc.out$mode.out,3)[1] #variance of intercept
  sig_beta <- tail(mcmc.out$mode.out,3)[2] #variance of site effect
  sig_gamma <- tail(mcmc.out$mode.out,3)[3] #variance of species effect
  }

data = data.frame(p_est, mse_shan, mbias_shan, sig_alpha, sig_beta, sig_gamma)
print(data)
```

